---
title: "Model Validation Z-tests"
author: "Jean-Luc Cartron, Matthieu Cartron, Hunter Thompson"
date: "2023-11-25"
output: github_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Model Validation Z-tests

## Requisite Packages

```{r}
library(readxl)
library(tidyverse)

```

## Read in Data

```{r}
Owl_data <- read_excel('D:\\Data\\Owl_Simulations_Final.xlsx')

```

## Variable Assignment


```{r}
LEOW_intersections <- Owl_data$Long_eared_Intersections
WESO_intersections <- Owl_data$Western_Screech_Intersections
WHSO_intersections <- Owl_data$Whiskered_Screech_Intersections
NSWO_intersections <- Owl_data$Saw_Whet_Intersections
PO_intersections <- Owl_data$Northern_Pygmy_Intersections
BOOW_intersections <- Owl_data$Boreal_Intersections
GHO_intersections <- Owl_data$Great_horned_Intersections
FO_intersections <- Owl_data$Flammulated_Intersections
```


## Z-tests

1. **Boreal Owl**

```{r}
BOOW_mean_prop <- mean(BOOW_intersections)/23
BOOW_mean_prop

prop.test(17, 23, p = BOOW_mean_prop, alternative = c("greater"), conf.level = 0.95, correct = TRUE)
```
2. **Whiskered Screech Owl**

```{r}
WHSO_mean_prop <- mean(WHSO_intersections)/112
WHSO_mean_prop

prop.test(71, 112, p = WHSO_mean_prop, alternative = c("greater"), conf.level = 0.95, correct = TRUE)
```

3. **Western Screech Owl**

```{r}
WESO_mean_prop <- mean(WESO_intersections)/137
WESO_mean_prop

prop.test(34, 137, p = WESO_mean_prop, alternative = c("greater"), conf.level = 0.95, correct = TRUE)
```

4. **Long Eared Owl**

```{r}
LEOW_mean_prop <- mean(LEOW_intersections)/23
LEOW_mean_prop

prop.test(14, 23, p = LEOW_mean_prop, alternative = c("greater"), conf.level = 0.95, correct = TRUE)
```

5. **Northern Saw-whet Owl**

```{r}
NSWO_mean_prop <- mean(NSWO_intersections)/34
NSWO_mean_prop

prop.test(22, 34, p = NSWO_mean_prop, alternative = c("greater"), conf.level = 0.95, correct = TRUE)
```
6. **Northern Pygmy Owl**

```{r}
PO_mean_prop <- mean(PO_intersections)/69
PO_mean_prop

prop.test(47, 69, p = PO_mean_prop, alternative = c("greater"), conf.level = 0.95, correct = TRUE)
```
7. **Flammulated Owl**

```{r}
FO_mean_prop <- mean(FO_intersections)/291
FO_mean_prop

prop.test(248, 291, p = FO_mean_prop, alternative = c("greater"), conf.level = 0.95, correct = TRUE)
```
8. **Greated horned owl**

```{r}
GHO_mean_prop <- (mean(GHO_intersections)/148) - 1e-10
GHO_mean_prop

prop.test(125, 148, p = GHO_mean_prop, alternative = c("greater"), conf.level = 0.95, correct = TRUE)

```
Note: Subtraction of arbitrarily small number from the Great horned owl proportion of intersections was made because prop.test() requires probabilities to be greater than zero and less than 1. The proportion of intersection should be 100%, and thus the p-value should be 1. Subtracting the arbitrarily small value allows for the test to be performed without changing the value of p.










